---
title: "NOCA Manuscript Code"

---


Required libraries
```{r}
library(AICcmodavg)
library(lme4)
library(MASS)
library(tidyverse)
library(performance)
library(MuMIn)
library(visreg) 
library(interactions)
library(ggplot2)
```


Dataset and scaling of independent variables
```{r}

NOCAspatial <- read.csv("~/NOCAspatial_short.csv",header=TRUE)

# Scaling variables
NOCAspatial$scShrubDtc <- c(scale(NOCAspatial$ShrubDtc, scale=TRUE))
NOCAspatial$scMidstoryDtc <- c(scale(NOCAspatial$MidstoryDtc, scale=TRUE))
NOCAspatial$scOverstoryDtc <- c(scale(NOCAspatial$OverstoryDtc, scale=TRUE))
NOCAspatial$scTotalGapDomin <- c(scale(NOCAspatial$totalGapDomin, scale=TRUE))
NOCAspatial$scNumGaps <- c(scale(NOCAspatial$numGaps, scale=TRUE))
NOCAspatial$scDisToEdge <- c(scale((NOCAspatial$DisToEdge), scale=TRUE))####### sqrt
NOCAspatial$scDeveloped_Open_Space <- c(scale(NOCAspatial$Developed_Open_Space, scale=TRUE))
NOCAspatial$scDeveloped_Low <- c(scale((NOCAspatial$Developed_Low), scale=TRUE))
NOCAspatial$scDeveloped_Med <- c(scale(NOCAspatial$Developed_Med, scale=TRUE))
NOCAspatial$scDeveloped_Hi <- c(scale(NOCAspatial$Developed_Hi, scale=TRUE))
NOCAspatial$scDeciduous_Forest <- c(scale(NOCAspatial$Deciduous_Forest, scale=TRUE))
NOCAspatial$scAreaStudySite <- c(scale(NOCAspatial$AreaStudySite, scale=TRUE))
NOCAspatial$NestHeight[is.na(NOCAspatial$NestHeight)]<-mean(NOCAspatial$NestHeight,na.rm=TRUE) #Getting rid of nest height NA's by estimation
NOCAspatial$scNestHeight <- c(scale(as.numeric(NOCAspatial$NestHeight), scale=TRUE))

colnames(NOCAspatial)
```


Exposure custom link function
```{r}
remove(..exposure) # Remove exposure before plotting
# logistic exposure function method 2. this one was used.         
#library(MASS)
logexp <- function(exposure = 1) {
  get_exposure <- function() {
    if (exists("..exposure", env=.GlobalEnv))
      return(get("..exposure", envir=.GlobalEnv))
    exposure
  }
  linkfun <- function(mu) qlogis(mu^(1/get_exposure()))
  linkinv <- function(eta) plogis(eta)^get_exposure()
  logit_mu_eta <- function(eta) {
    ifelse(abs(eta)>30,.Machine$double.eps,
           exp(eta)/(1+exp(eta))^2)
  }
  mu.eta <- function(eta) {       
    get_exposure() * plogis(eta)^(get_exposure()-1) *
      logit_mu_eta(eta)
  }
  valideta <- function(eta) TRUE
  link <- paste("logexp(", deparse(substitute(exposure)), ")",
                sep="")
  structure(list(linkfun = linkfun, linkinv = linkinv,
                 mu.eta = mu.eta, valideta = valideta, 
                 name = link),
            class = "link-glm")
}
```


Candidate models
```{r}
Cand <- list()

### NULL MODEL
Cand[[1]] <- lme4::glmer(Success ~ + (1|Year:Site), data = NOCAspatial, family = binomial(link=logexp(NOCAspatial$Exposure)), 
                         control = glmerControl(optimizer = "bobyqa"),nAGQ = 4); summary(Cand[[1]])

####### NEST-SCALE MODELS

# OVERSTORY DENSITY (Dtc = Detections)
Cand[[2]] <- lme4::glmer(Success ~ scOverstoryDtc + (1|Year:Site), data = NOCAspatial, family = binomial(link=logexp(NOCAspatial$Exposure)), 
                         control = glmerControl(optimizer = "bobyqa"), nAGQ = 4); summary(Cand[[2]]) 

# SHRUB DENSITY 
Cand[[3]] <- lme4::glmer(Success ~ scShrubDtc + (1|Year:Site),data = NOCAspatial, family = binomial(link=logexp(NOCAspatial$Exposure)), 
                         control = glmerControl(optimizer = "bobyqa"),nAGQ = 4); summary(Cand[[3]])

# MIDSTORY DENSITY
Cand[[4]] <- lme4::glmer(Success ~ scMidstoryDtc + (1|Year:Site),data = NOCAspatial, family = binomial(link=logexp(NOCAspatial$Exposure)), 
                         control = glmerControl(optimizer = "bobyqa"),nAGQ = 4); summary(Cand[[4]])

## ADDED EFFECTS OF OVERSTORY AND SHRUB
Cand[[5]] <- lme4::glmer(Success ~ scOverstoryDtc + scShrubDtc + (1|Year:Site), data = NOCAspatial, family = binomial(link=logexp(NOCAspatial$Exposure)), 
                         control = glmerControl(optimizer = "bobyqa"), nAGQ = 4); summary(Cand[[5]]) 

## ADDED EFFECTS OF OVERSTORY AND MIDSTORY
Cand[[6]] <- lme4::glmer(Success ~ scOverstoryDtc + scMidstoryDtc + (1|Year:Site), data = NOCAspatial, family = binomial(link=logexp(NOCAspatial$Exposure)), 
                         control = glmerControl(optimizer = "bobyqa"), nAGQ = 4); summary(Cand[[6]]) 

# GAP NUMBER
Cand[[7]] <- lme4::glmer(Success ~ scNumGaps + (1|Year:Site),data = NOCAspatial, family = binomial(link=logexp(NOCAspatial$Exposure)), 
                         control = glmerControl(optimizer = "bobyqa"),nAGQ = 4); summary(Cand[[7]])

# GAP DOMINANCE (Gap weight)
Cand[[8]] <- lme4::glmer(Success ~ scTotalGapDomin + (1|Year:Site),data = NOCAspatial, family = binomial(link=logexp(NOCAspatial$Exposure)), 
                         control = glmerControl(optimizer = "bobyqa"),nAGQ = 4); summary(Cand[[8]])

# NEST HEIGHT
Cand[[9]] <- lme4::glmer(Success ~ scNestHeight + (1|Year:Site), data = NOCAspatial, family = binomial(link=logexp(NOCAspatial$Exposure)), 
                         control = glmerControl(optimizer = "bobyqa"),nAGQ = 4); summary(Cand[[9]])



### MULTIPLE NEST-SCALE, RETAINING OVERSTORY DENSITY AND GAP VARIABLES

# OVERSTORY + NUMBER OF GAPS
Cand[[10]] <- lme4::glmer(Success ~ scOverstoryDtc + scNumGaps + (1|Year:Site), data = NOCAspatial, family = binomial(link=logexp(NOCAspatial$Exposure)), 
                         control = glmerControl(optimizer = "bobyqa"),nAGQ = 4); summary(Cand[[10]]) 
 performance::check_collinearity(Cand[[10]]) # Low VIF 

# OVERSTORY + GAP DOMIN
Cand[[11]] <- lme4::glmer(Success ~ scOverstoryDtc + scTotalGapDomin + (1|Year:Site), data = NOCAspatial, family = binomial(link=logexp(NOCAspatial$Exposure)), 
                         control = glmerControl(optimizer = "bobyqa"),nAGQ = 4); summary(Cand[[11]]) 

# OVERSTORY + NEST HEIGHT
Cand[[12]] <- lme4::glmer(Success ~ scOverstoryDtc + scNestHeight + (1|Year:Site), data = NOCAspatial, family = binomial(link=logexp(NOCAspatial$Exposure)), 
                          control = glmerControl(optimizer = "bobyqa"),nAGQ = 4); summary(Cand[[12]])

# GAP NUMBER + NEST HEIGHT
Cand[[13]] <- lme4::glmer(Success ~ scNumGaps + scNestHeight + (1|Year:Site), data = NOCAspatial, family = binomial(link=logexp(NOCAspatial$Exposure)),
                          control = glmerControl(optimizer = "bobyqa"),nAGQ = 4); summary(Cand[[13]])


# LANDSCAPE AND PATCH VARIABLES ADDED TO THE BEST NEST SITE-LEVEL MODELS

Cand[[14]] <- lme4::glmer(Success ~ scOverstoryDtc + scNumGaps + scAreaStudySite + (1|Year:Site), data = NOCAspatial, family = binomial(link=logexp(NOCAspatial$Exposure)), 
                          control = glmerControl(optimizer = "bobyqa"),nAGQ = 4); summary(Cand[[14]]) 

Cand[[15]] <- lme4::glmer(Success ~ scOverstoryDtc + scNumGaps + scDisToEdge + (1|Year:Site), data = NOCAspatial, family = binomial(link=logexp(NOCAspatial$Exposure)), 
                          control = glmerControl(optimizer = "bobyqa"),nAGQ = 4); summary(Cand[[15]]) 

Cand[[16]] <- lme4::glmer(Success ~ scOverstoryDtc + scNumGaps +  scDeveloped_Low + (1|Year:Site), data = NOCAspatial, family = binomial(link=logexp(NOCAspatial$Exposure)), 
                          control = glmerControl(optimizer = "bobyqa"),nAGQ = 4); summary(Cand[[16]]) 

Cand[[17]] <- lme4::glmer(Success ~ scOverstoryDtc + scNumGaps +  scDeciduous_Forest + (1|Year:Site), data = NOCAspatial, family = binomial(link=logexp(NOCAspatial$Exposure)), 
                          control = glmerControl(optimizer = "bobyqa"),nAGQ = 4); summary(Cand[[17]]) 



Cand[[18]] <- lme4::glmer(Success ~ scOverstoryDtc + scNumGaps + scDeveloped_Low + scDisToEdge + (1|Year:Site), data = NOCAspatial, family = binomial(link=logexp(NOCAspatial$Exposure)),
                          control = glmerControl(optimizer = "bobyqa"),nAGQ = 4); summary(Cand[[18]]) 

Cand[[19]] <- lme4::glmer(Success ~ scOverstoryDtc + scNumGaps +  scDeciduous_Forest + scDisToEdge + (1|Year:Site), data = NOCAspatial, family = binomial(link=logexp(NOCAspatial$Exposure)), 
                          control = glmerControl(optimizer = "bobyqa"),nAGQ = 4); summary(Cand[[19]]) 


# ECOLOGICALLY INFORMED/INTERPRETABLE INTERACTIONS

Cand[[20]] <- lme4::glmer(Success ~ scOverstoryDtc + scNumGaps + scDeveloped_Low * scDisToEdge + (1|Year:Site), data = NOCAspatial, family = binomial(link=logexp(NOCAspatial$Exposure)),
                          control = glmerControl(optimizer = "bobyqa"),nAGQ = 4); summary(Cand[[20]]) 

Cand[[21]] <- lme4::glmer(Success ~ scOverstoryDtc + scNumGaps*scDeveloped_Low + scDisToEdge + (1|Year:Site), data = NOCAspatial, family = binomial(link=logexp(NOCAspatial$Exposure)), 
                          control = glmerControl(optimizer = "bobyqa"),nAGQ = 4); summary( Cand[[21]]) 

Cand[[22]] <- lme4::glmer(Success ~ scOverstoryDtc + scNumGaps*scDisToEdge + scDeveloped_Low  + (1|Year:Site), data = NOCAspatial, family = binomial(link=logexp(NOCAspatial$Exposure)), 
                          control = glmerControl(optimizer = "bobyqa"),nAGQ = 4); summary( Cand[[22]]) 

Cand[[23]] <- lme4::glmer(Success ~ scOverstoryDtc*scDisToEdge + scNumGaps + scDeveloped_Low  + (1|Year:Site), data = NOCAspatial, family = binomial(link=logexp(NOCAspatial$Exposure)), 
                          control = glmerControl(optimizer = "bobyqa"),nAGQ = 4); summary( Cand[[23]]) 

```


AIC table and top model correlations and convergence
```{r}  

# Make AIC table
  AICtable <-   aictab(Cand, sort = TRUE); AICtable
  AICt<-as.data.frame(AICtable)
  write_csv(AICt, "~/Downloads/AICtable.csv")

# Checking variance inflation factors
  performance::check_collinearity(Cand[[20]]) # Low correlation
  
# Checking convergence 
  performance::check_convergence(Cand[[20]]) # True

```


Plotting main effects using visreg
```{r}
# Set ..exposure variable in the global environment,
..exposure <- mean(NOCAspatial$Exposure)

# Marginal plots, and see fitted values
  a<-visreg(Cand[[27]], "scOverstoryDtc", scale="response" ,rug=2, band=F); #a$fit
  b<-visreg(Cand[[27]], "scNumGaps", scale="response" ,rug=2, band=F);# b$fit
  c<-visreg(Cand[[27]], "scDeveloped_Low", scale="response" ,rug=2); c$fit
  d<-visreg(Cand[[27]], "scDisToEdge", scale="response" ,rug=2); d$fit
  e<-visreg(Cand[[27]], "scDisToEdge", by="scDeveloped_Low", overlay=T, type="contrast", breaks=2); e$fit #Doesn't average across sites, not ideal.

#Confirmation - Year-only random effect model resolves visreg convergence issues, and results are very similar.
fit2 <- lme4::glmer(Success ~ scOverstoryDtc + scNumGaps + scDeveloped_Low * scDisToEdge + (1|Year), data = NOCAspatial, family = binomial(link=logexp(NOCAspatial$Exposure)), 
                          control = glmerControl(optimizer = "bobyqa"),nAGQ = 4); summary(fit2) 
```


Interactive effect plot using interact_plot (visreg convergene issues avoided, more featuers)
```{r}

#Based on 4 most and 4 least developed (more interpretable; in publication)

NOCAspatial$Urban <- cut(NOCAspatial$scDeveloped_Low,
                         breaks=c(-5, 0.23, 5),
                         labels=c('Less', 'More'))

fit4 <- lme4::glmer(Success ~ scOverstoryDtc + scNumGaps + Urban * scDisToEdge + (1|Year), data = NOCAspatial, family = binomial(link=logexp(NOCAspatial$Exposure)), 
                          control = glmerControl(optimizer = "bobyqa"),nAGQ = 4)

interactions::interact_plot(fit4, pred=scDisToEdge, modx=Urban, colors=c("cornflowerblue","black"), int.width = .95,
                            lty=1,  plot.points=T, point.shape=T,#modx.values="plus-minus",
                            x.label="\n Distance to Edge", y.label = "Probability of Success \n",
                           legend.label = "Development\n Level", modx.labels = c("Less Developed", "More Developed"),
                            point.size = 3, line.thickness = 1.5, vary.lty = T, jitter=0.02)+
#scale_color_gradient(low = "rosybrown1", high = "red4")+
theme_bw()+
  theme(panel.grid=element_blank()) 

######

# Based on mean and sd's (intersection more accurately depicted; not in publication)

interactions::interact_plot(Cand[[20]], pred=scDisToEdge, modx=scDeveloped_Low, colors=c("cornflowerblue","black"),interval = TRUE,int.type = "prediction", int.width = .95,
                            lty=1,  plot.points=T, point.shape=T, modx.values = c(-1, 1), #Set at +-1 sd or modx.values="plus-minus",
                            x.label="\n Distance to Edge", y.label = "Probability of Success \n",
                            legend.label = "Development\n Level", modx.labels = c("Less Developed", "More Developed"),
                            point.size = 4, line.thickness = 1.5, vary.lty = T, jitter=0.01)+ 
 scale_color_gradient(low = "cornflowerblue", high = "black")+
  theme_bw()+
  theme(panel.grid=element_blank()) #+
#  geom_vline(xintercept = 0.8, linetype = "dotted", color = "blue", size = 0.2)
```


Producing a correlation matrix plot
```{r}
library(corrplot)
colnames(NOCAspatial)
corNOCAspatial <- NOCAspatial[, c(24:36)] # Select the columns of interest

## Change column names to match manuscript text
colnames(corNOCAspatial) <- c("ShrubDen", "MidstoryDen","OSden","GapWeight", "NumGaps", 
                              "DisEdge", "DevelopedOpen", "Develop","DevMed", "DevHi", "DecForest",
                              "PatchArea", "NestHeight")

cor <- cor(corNOCAspatial)

corrplot(cor, type="upper", order="hclust", tl.cex = 0.9, cl.cex = 1, mar=c(0,0,0,0), tl.col='black') ## mar is margins (set to zero on all sides)

psych::describe(NOCAspatial$Deciduous_Forest)


```


Among-site distances
```{r}
AmongSite <- c(9500	,1390	,3970	,6980,	6120	,5210	,5630,
               10050	,13540,	16320,	15610	,10770,	15090,
               4140,	6250	,5640,	6710,	5100,
               4360,	2690	,6140	,1650,
               1800,	10690,	4460,
               8890,	2530,
               6780)

```

